### 2019-09-17 hw3/MSimonsHW3.tex
Starting hw3 problem 5

\documentclass{article}
\title{Homework 3}
\author{Madilyn Simons}
\date{}
\setlength{\parindent}{0pt}
\pdfpagewidth 8.5in
\pdfpageheight 11in

\usepackage{amssymb}
\usepackage{amsmath}

\begin{document}
\maketitle

\begin{enumerate}

\item TODO

\item TODO

\item TODO

\item TODO

\item Assume $M$ is a unit.  Therefore there exists some $M^{-1} \in M_2(\mathbb{Z})$
such that $MM^{-1} = I_2$.  We find $M^{-1}$ using Gaussian Elimination:
\[
\left[
\begin{array}{cc|cc}
a & b & 1 & 0 \\
c & d & 0 & 1
\end{array}
\right]
\Leftrightarrow
\left[
\begin{array}{cc|cc}
a & b & 1 & 0 \\
0 & \frac{ad-bc}{a} & \frac{-c}{a} & 1
\end{array}
\right]
\]
\[
\Leftrightarrow
\left[
\begin{array}{cc|cc}
a & b & 1 & 0 \\
0 & 1 & \frac{-c}{ad-bc} & \frac{a}{ad-bc}
\end{array}
\right]
\Leftrightarrow
\left[
\begin{array}{cc|cc}
a & 0 & \frac{ad}{ad-bc} & \frac{-ab}{ad-bc} \\
0 & 1 & \frac{-c}{ad-bc} & \frac{a}{ad-bc}
\end{array}
\right]
\]
\[
\Leftrightarrow
\left[
\begin{array}{cc|cc}
1 & 0 & \frac{d}{ad-bc} & \frac{-b}{ad-bc} \\
0 & 1 & \frac{-c}{ad-bc} & \frac{a}{ad-bc}
\end{array}
\right]
\]

Therefore $M^{-1} = \begin{bmatrix} \frac{d}{ad-bc} & \frac{-b}{ad-bc} \\
\frac{-c}{ad-bc} & \frac{a}{ad-bc}\end{bmatrix}$.  Since $M^{-1} \in M_2(\mathbb{Z})$,
each of its elements must be integers.  Its elements can only be
integers if $ad-bc$ evenly divides $a$, $b$, $c$, and $d$. \\
\\
First, assume $a$, $b$, $c$, and $d$ are relatively prime.  If so, then the
only integers that evenly divide all four are $\pm1$.  Therefore
${ad-bc}$ must be $\pm1$. \\
\\
Otherwise, $a$, $b$, $c$, and $d$ are not relatively prime.  Let $m$ be the
greatest common factor of the $a$, $b$, $c$, and $d$.  Now, let $a = mx$,
$b = my$, $c = mz$ and $d = mw$ for some integers $x$, $y$, $z$, and $w$.
As such, $ad - bc = m^2xw - m^yz = m^2(xw - yz)$.  This means that
$m^2$ and $(xw - yz)$ must also be common factors of $a$, $b$, $c$, and $d$.
If $m$ is the greatest common factor, then $m^2 \leq \vertm\vert$.  This is only
possible if $m = 1$.  Likewise, $\vert(xw - yz)\vert \leq \vertm\vert$.
Since $m = 1$, $(xw - yz)$ must be $\pm1$ or $0$, but it can't be $0$ since
$m^2 \cdot 0 = 0$, which can't divide any integer.  Since $m^2 = 1$ and
$xw - yz = \pm1$, $ad - bc = \pm1$.

\end{enumerate}
\end{document}