### 2019-09-19 hw3/MSimonsHW3.tex
Finished number 5 in hw 3

\documentclass{article}
\title{Homework 3}
\author{Madilyn Simons}
\date{}
\setlength{\parindent}{0pt}
\pdfpagewidth 8.5in
\pdfpageheight 11in

\usepackage{amssymb}
\usepackage{amsmath}

\begin{document}
\maketitle

\begin{enumerate}

\item TODO

\item TODO

\item TODO

\item TODO

\item Assume $M$ is a unit.  Therefore there exists some $M^{-1} \in M_2(\mathbb{Z})$
such that $MM^{-1} = I_2$.  We find $M^{-1}$ using Gaussian Elimination:
\[
\left[
\begin{array}{cc|cc}
a & b & 1 & 0 \\
c & d & 0 & 1
\end{array}
\right]
\Leftrightarrow
\left[
\begin{array}{cc|cc}
a & b & 1 & 0 \\
0 & \frac{ad-bc}{a} & \frac{-c}{a} & 1
\end{array}
\right]
\]
\[
\Leftrightarrow
\left[
\begin{array}{cc|cc}
a & b & 1 & 0 \\
0 & 1 & \frac{-c}{ad-bc} & \frac{a}{ad-bc}
\end{array}
\right]
\Leftrightarrow
\left[
\begin{array}{cc|cc}
a & 0 & \frac{ad}{ad-bc} & \frac{-ab}{ad-bc} \\
0 & 1 & \frac{-c}{ad-bc} & \frac{a}{ad-bc}
\end{array}
\right]
\]
\[
\Leftrightarrow
\left[
\begin{array}{cc|cc}
1 & 0 & \frac{d}{ad-bc} & \frac{-b}{ad-bc} \\
0 & 1 & \frac{-c}{ad-bc} & \frac{a}{ad-bc}
\end{array}
\right]
\]

Therefore $M^{-1} = \begin{bmatrix} \frac{d}{ad-bc} & \frac{-b}{ad-bc} \\
\frac{-c}{ad-bc} & \frac{a}{ad-bc}\end{bmatrix}$.  Since $M^{-1} \in M_2(\mathbb{Z})$,
each of its elements must be integers.  Its elements can only be
integers if $ad-bc$ evenly divides $a$, $b$, $c$, and $d$. \\
\\
Let $m$ be the greatest common denominator of $a$, $b$, $c$, and $d$.
Let $a = mx$, $b = my$, $c = mz$, and $d = mw$ for some integers
$x$, $y$, $z$, and $w$ that are not divisible by $m$.
As such $ad-bc = mxmw - mymz = m^2(xw - yz)$.  Since $m$ is the greatest common
denominator and $m^2$ is also a common denominator, $m^2 \leq m$ and this is
only possible if $m = 1$.  Therefore, the greatest common denominator of
$a$, $b$, $c$, and $d$ is 1.  If this is the case, $ad-bc$ must be $\pm1$ since
only $\pm1$ can evenly $a$, $b$, $c$, and $d$.
\\
\\
Next, assume $ad - bc = \pm1$.  By the last proof, $M$ can only be a unit if
$MM^{-1} = I_2$ and $M^{-1} \in M_2(\mathbb{Z})$.  If $ad - bc = 1$, then
$M^{-1} = \begin{bmatrix} d & -b \\ -c & a \end{bmatrix}$ and $MM^{-1} = I_2$.
If $ad - bc = -1$, then $M^{-1} = \begin{bmatrix} -d & b \\ c & -a \end{bmatrix}$
and $MM^{-1} = I_2$.
\\
\\
Thus, $M \in M_2(\mathbb{Z})$ is a unit if and only if $ad - bc = \pm1$.

\end{enumerate}
\end{document}