### 2019-09-26 hw4/MSimonsHW4.tex
Starting hw4

\documentclass{article}
\title{Homework 4}
\author{Madilyn Simons}
\date{}
\setlength{\parindent}{0pt}
\pdfpagewidth 8.5in
\pdfpageheight 11in

\usepackage{amssymb}
\usepackage{amsmath}

\begin{document}
\maketitle

\begin{enumerate}

\item TODO

\item TODO

\item TODO

\item TODO

\item TODO

\item TODO

\end{enumerate}
\end{document}