### 2019-10-01 hw5/MSimonsHW5.tex
HW5 question 1

\documentclass{article}
\title{Homework 5}
\author{Madilyn Simons}
\date{}
\setlength{\parindent}{0pt}
\pdfpagewidth 8.5in
\pdfpageheight 11in

\usepackage{amssymb}
\usepackage{amsmath}

\begin{document}
\maketitle

\begin{enumerate}

\item To prove $\varphi_{\alpha}$ is closed under addition, first let $f$, $g$
be any elements of $S[x]$.  We know that $f+g$ is also an element of
$S[x]$ because the coefficients of $f$ and $g$ are elements of $S$, which is a
ring.  This implies that the coefficients of $f$ and $g$ are closed under
addition, so the coefficients of $f+g$ are also in $S$.
\\ \\
With this having be said, we can prove $\varphi_{\alpha}$ is closed under addition
as such:
\[
\varphi_{\alpha}(f+g) = (f+g)(\alpha) = f(\alpha) + g(\alpha)
= \varphi_{\alpha}(f) + \varphi_{\alpha}(g)
\]
We know $f \cdot g$ is an element of $S[x]$ since $S$ is closed under multiplication,
so the coefficients of $f \cdot g$ are elements of $S$.
\\ \\
We can prove $\varphi_{\alpha}$ is closed under multiplication as such:
\[
\varphi_{\alpha}(f \cdot g) = (f \cdot g)(\alpha) = f(\alpha) \cdot g(\alpha)
= \varphi_{\alpha}(f) \cdot \varphi_{\alpha}(g)
\]
Thus, $\varphi_{\alpha}$ is a ring homomorphism.

\item TODO
\item TODO

\end{enumerate}
\end{document}