### 2019-10-01 hw5/MSimonsHW5.tex
HW 5 number 2

\documentclass{article}
\title{Homework 5}
\author{Madilyn Simons}
\date{}
\setlength{\parindent}{0pt}
\pdfpagewidth 8.5in
\pdfpageheight 11in

\usepackage{amssymb}
\usepackage{amsmath}

\begin{document}
\maketitle

\begin{enumerate}

\item To prove $\varphi_{\alpha}$ is closed under addition, first let $f$, $g$
be any elements of $S[x]$.  We know that $f+g$ is also an element of
$S[x]$ because the coefficients of $f$ and $g$ are elements of $S$, which is a
ring.  This implies that the coefficients of $f$ and $g$ are closed under
addition, so the coefficients of $f+g$ are also in $S$.
\\ \\
With this having be said, we can prove $\varphi_{\alpha}$ is closed under addition
as such:
\[
\varphi_{\alpha}(f+g) = (f+g)(\alpha) = f(\alpha) + g(\alpha)
= \varphi_{\alpha}(f) + \varphi_{\alpha}(g)
\]
We know $f \cdot g$ is an element of $S[x]$ since $S$ is closed under multiplication,
so the coefficients of $f \cdot g$ are elements of $S$.
\\ \\
We can prove $\varphi_{\alpha}$ is closed under multiplication as such:
\[
\varphi_{\alpha}(f \cdot g) = (f \cdot g)(\alpha) = f(\alpha) \cdot g(\alpha)
= \varphi_{\alpha}(f) \cdot \varphi_{\alpha}(g)
\]
Thus, $\varphi_{\alpha}$ is a ring homomorphism.

\item Let $f(\alpha)$, $g(\alpha)$ be any elements in $S[\alpha]$.  By the
definition of $S[\alpha]$, $f(x)$, $g(x)$ as elements of $S[x]$.  Since $S[x]$
is a ring homomorphism, $(f+g)(x)$ is an element of $S[x]$.  If $(f+g)(x)$ is an
element of $S[x]$, then $(f+g)(\alpha)$ is an element of $S[\alpha]$ and
$S[\alpha]$ is closed under addition.
\\
\\
Similarly, $f(x)g(x)$ is an element of $S[x]$.  This implies that
$f(\alpha)g(\alpha)$ is an element of $S[\alpha]$ and so $S[\alpha]$ is
closed under multiplication.
\\
\\
For $S[\alpha]$ to be a subring of $R$, $0_R$ must be an element of $S[\alpha]$.
Define the function $f:S[x] \rightarrow R$ by $f(x) = 0_R$.  As such,
$f(\alpha) = 0_R$ is an element of $S[\alpha]$.
\\
\\
Let $f(x) = a_0 + a_{1}x + ... + a_{n}x^n$ be any element of $S[x]$.
By definition, $a_i$ is an element of $R$ for all non-negative integers $i$, and
for all $a_i$, there exists a $b_i$ in $R$ such that $a_i + b_i = 0_R$.
Therefore, there exists a function $g(x) = b_0 + b_{1}x + ... + b_{n}x^n$ in
$S[x]$ such that $f(x) + g(x) = 0_R$.  As a direct result, for any $f(\alpha)$
in $S[\alpha]$ there exists a $g(\alpha)$ such that
$f(\alpha) + g(\alpha) = 0_R$.
\\
\\
Thus, $S[\alpha]$ is a subring of $R$.

\item TODO

\end{enumerate}
\end{document}