### 2019-10-02 hw5/MSimonsHW5.tex
corrected HW 5 problem 2

\documentclass{article}
\title{Homework 5}
\author{Madilyn Simons}
\date{}
\setlength{\parindent}{0pt}
\pdfpagewidth 8.5in
\pdfpageheight 11in

\usepackage{amssymb}
\usepackage{amsmath}

\begin{document}
\maketitle

\begin{enumerate}

\item Assume that if $f(x)$, $g(x)$, $s(x)$, $d(x)$, $p(x)$ in $S[x]$ satisfy
\[
f(x) + g(x) = s(x)
\] \[
f(x) - g(x) = d(x)
\] \[
f(x)g(x) = p(x)
\]
then, for any $\alpha$ in $R$,
\[
\] \[
f(\alpha) + g(\alpha) = s(\alpha)
\] \[
f(\alpha) - g(\alpha) = d(\alpha)
\] \[
f(\alpha)g(\alpha) = p(\alpha)
\]
We can prove that $\phi_{\alpha}$ is closed under addition as such:
\[
\phi_{\alpha}(f) + \phi_{\alpha}(g) = f(\alpha) + g(\alpha) = s(\alpha) = \phi_{\alpha}(s)
\] \[
\phi_{\alpha}(f) - \phi_{\alpha}(g) = f(\alpha) - g(\alpha) = d(\alpha) = \phi_{\alpha}(d)
\]
We can prove that $\phi_{\alpha}$ is closed under multiplication as such:
\[
\phi_{\alpha}(f)\phi_{\alpha}(g) = f(\alpha)g(\alpha) = p(\alpha) = \phi_{\alpha}(p)
\]
Thus, $\phi_{\alpha}$ is a ring homomorphism.

\item Let $f(\alpha)$, $g(\alpha)$ be elements in $S[\alpha]$.  This implies
that $f(x)$ and $g(x)$ are elements of $S[x]$.  Since $S$ is a ring, its
elements are closed under addition, implying that $S[x]$ is closed under
addition as well.  Let $f(x) + g(x) = s(x)$ and $f(x) - g(x) = d(x)$.  Since
$s(x)$ and $d(x)$ are in $S[x]$, $s(\alpha) = f(\alpha) + g(\alpha)$ and
$d(\alpha) = f(\alpha) - g(\alpha)$ are elements of $S[\alpha]$ and $S[\alpha]$
is closed under addition.
\\
\\
Similarly, since $S$ is closed under multiplication, we can assume $p(x) = f(x)g(x)$
is an element of $S[x]$.  Therefore $p(\alpha) = f(\alpha)g(\alpha)$ is an
element of $S[\alpha]$ and $S[alpha]$ is closed under multiplication.
\\
\\
Thus, $S[\alpha]$ is a subring of $R$.

\item Let $(a_0 + a_{1}x + ... + a_{n}x^n)$, $(b_0 + b_{1}x + ... + b_{n}x^n)$
be elements of $R_{1}[x]$.
\\
\\
We can prove $G$ is closed under addition as such:
\[
G((a_0 + a_{1}x + ... + a_{n}x^{n}) + (b_0 + b_{1}x + ... + b_{n}x^{n}))
\] \[
= G(a_0 + a_{1}x + ... + a_{n}x^n + b_0 + b_{1}x + ... + b_{n}x^{n})
\] \[
= F(a_{0}) + F(a_{1})x + ... + F(a_{n})x^{n} + F(b_{0}) + F(b_{1})x + ... + F(b_{n})x^{n}
\] \[
= (F(a_{0}) + F(a_{1})x + ... + F(a_{n})x^{n}) + (F(b_{0}) + F(b_{1})x + ... + F(b_{n})x^{n})
\] \[
= G(a_0 + a_{1}x + ... + a_{n}x^{n}) + G(b_0 + b_{1}x + ... + b_{n}x^{n})
\]
We can prove $G$ is closed under multiplication as such:
\[
G((a_0 + a_{1}x + ... + a_{n}x^{n}) (b_0 + b_{1}x + ... + b_{n}x^{n}))
\] \[
= G (a_{0}b_{0} + ... + a_{n}b_{m}x^{n+m})
\] \[
= F(a_{0}b_{0}) + ... + F(a_{n}b_{m})x^{n+m}
\] \[
= F(a_{0})F(b_{0}) + ... + F(a_{n})F(b_{m})x^{n+m}
\] \[
= (F(a_{0}) + F(a_{1})x + ... + F(a_{n})x^{n})(F(b_{0}) + F(b_{1})x + ... + F(b_{m})x^{m})
\] \[
= G(a_0 + a_{1}x + ... + a_{n}x^{n}) G(b_0 + b_{1}x + ... + b_{n}x^{n})
\]
Therefore $G$ is also a a ring homomorphism.

\end{enumerate}
\end{document}