### 2019-10-30 hw7/MSimonsHW7.tex
added to hw7 problem 2

\documentclass{article}
\usepackage{amsmath,amssymb}
\newcommand{\Z}{\mathbb{Z}}
\title{Homework 7}
\author{Madilyn Simons}
\date{}
\setlength{\parindent}{0pt}
\pdfpagewidth 8.5in
\pdfpageheight 11in

\usepackage{amssymb}
\usepackage{amsmath}

\begin{document}
\maketitle

\begin{enumerate}

\item TODO

\item Let $f(x) = x^5 - 4x^4 + 4d^3 + 3x^2 -26x + 21$.
If $p = 2$, then $\overline{f}(x) = x^5 + x^2 + 1$ in $\mathbb{Z}_{p}[x]$.
Since $\overline{f}(0) = \overline{f}(1) = 1$, $\overline{f}(x)$ is
irreducible in $\mathbb{Z}_{p}[x]$.  Therefore, by Theorem 4.25, $f(x)$ is
irreducible in $\mathbb{Q}[x]$.

\item TODO

\end{enumerate}
\end{document}