### 2019-11-29 hw9/MSimonsHW9.tex
hw9 problems 1ab

\documentclass{article}
\usepackage{amsmath,amssymb}
\newcommand{\Z}{\mathbb{Z}}
\title{Homework 8}
\author{Madilyn Simons}
\date{}
\setlength{\parindent}{0pt}
\pdfpagewidth 8.5in
\pdfpageheight 11in

\usepackage{amssymb}
\usepackage{amsmath}

\begin{document}
\maketitle

\begin{enumerate}

\item
\begin{enumerate}

\item Let $I$ be a nonempty ideal in $F$ such that $I$ does not equal $F$.
Let $x \in F$ and $x \notin I$, let $a \in I$, and let $a^{-1}$ be the
multiplicative inverse of $a$.  If $a \in I$, then $ax \in I$ and
$a^{-1}ax \in I$.  However, $a^{-1}ax = 1_{F}x = x$ and $x \notin I$ so this
is a contradiction.

\item Since $f$ is a homomorphism of rings, the kernel of $f$ is an ideal
of $F$.  The only ideals of $F$ are $F$ and $(0)$.  If the kernel of
$f$ is $(0)$, then $f$ is injective.  If the kernel of $f$ is $F$, then
$f$ is the zero function.

\end{enumerate}

\end{enumerate}
\end{document}